---
title: "Econometrics Homework 1"
author: "Charles Reid"
date: "9/14/2020"
---


## What are the names of the people in your study group?

* Nikki
* Chistopher
* Musbaou
* Isabella
* Mostafa
* Charles

## Do about 20 experiments where you roll a dice and record whether the result is a 6 or not.

```{r}
Dice_Experiment <- sample(x = 1:6, size = 20, replace = TRUE)

Dice_Experiment

sum(Dice_Experiment == 6)

plot(table(Dice_Experiment), xlab = 'Dice number', ylab = 'Frequency', main = '20 Rolls of 2 Fair Dice')

```

## “Adjust” the dice. Can you get them to roll a 6 more or less often than would be expected? How would you know - what is “more often”? Can you write some code in R that will simulate a fair roll?
```{r}
Unfair_Dice_Experiment <- sample(x = 1:6, size = 20, replace = TRUE, prob = c(0.2, 0.1, 0.1, 0.05, 0.05, 0.5))

Unfair_Dice_Experiment

sum(Unfair_Dice_Experiment == 6)

plot(table(Unfair_Dice_Experiment), xlab = 'Dice number', ylab = 'Frequency', main = '20 Rolls of 2 VERY Un-Fair Dice')
```

## Replicate the commands given in the lecture notes R Basics for Lecture 1. 
```{r}
load("acs2017_ny_data.RData")
acs2017_ny[1:10,1:7]
```


```{r}
attach(acs2017_ny)
```


```{r}
summary(acs2017_ny)
```
196585 people are in the data set.
```{r}
print(NN_obs <- length(AGE))
```


## Do some simple stats on the PUMS-NY data:
Summary of Females
```{r}
summary(AGE[female ==1])
```
Summary of Non-females.
```{r}
summary(AGE[!female])
```

Those notes request that you find average ages for: 
men 
```{r}
mean(AGE[!female == 1])
```


and women after accounting for the top-coding. 
```{r}
mean(AGE[female == 1])
```


Tell me something else interesting, that you learned from the data: 

What is surpizing is that women tend to be a little older, about two years older than men.
```{r}
summary(EDUC[female == 1])
summary(EDUC[!female])
```

Both genders have a mean education of about 15:
```{r}
mean(educ_college [female == 1])*100
mean(educ_college [!female])*100
```


For example about educational attainments in different neighborhoods in the city. Are there surprises for you?:
```{r}
summary(EDUC[in_Brooklyn])
summary(EDUC[in_Queens])
summary(EDUC[in_Bronx])
summary(EDUC[in_Manhattan])
summary(EDUC[in_Westchester])
```
People in Brooklyn have the highest level of education.

Differences in means can be complicated. Find the mean return on SP500 index (choose a time period). What is the mean return on days when the previous day’s return was positive? When the previous 2 days were positive? Negative? Now read about “hot hands fallacy” and tell if you think that helps investment strategy. (You might start with this tweet, and read the papers referenced.)
```{r}
options(readr.num_columns = 0)
library(readr)
S_P_Stock_Price <- read_csv("all_stocks_5yr.csv")

head(S_P_Stock_Price)
```

Adj close numbers
```{r}
S_P_Stock_Price[1:30, "close"]
```
Isolated closing price and
```{r}
sp_pricedate <-S_P_Stock_Price[, "date", drop = FALSE]
sp_pricenew <-S_P_Stock_Price[, "close", drop = FALSE]
cbind(sp_pricedate,sp_pricenew)
```
Calculating Close Returns
```{r}
n <-nrow(sp_pricenew)
sp_returns <- log(sp_pricenew[2:n,1])-log(sp_pricenew[1:(n-1),1])
returns <- sp_returns*100
summary(returns)
```

```{r}
xdate <- sp_pricedate[2:619040,"date"]
df1 <- cbind.data.frame(xdate,returns)
print(df1)
```
A hot-hand fallacy occurs when gamblers think that a winning streak is more likely to continue. 






